---
title: 
  \vspace{3in} "Visualizing, Modeling, and Forecasting Crashes in Wake County, North Carolina"
author: 
  - Cortlyn Morris, Kelly Wenzlof, Ayan Gulzar
  - Research Mentors Dr. Sayed Mostafa and Chinedu Jude Nzekwe 
date: "7/14/2021"
header-includes:
    - \usepackage{setspace}\doublespacing
output: 
  pdf_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
\tableofcontents
\newpage


# Abstract

Studies have shown that road traffic crashes are one of the leading causes of non-natural death for U.S citizens. There are many different ways that road crashes occur, and several various factors lead to it. It is essential to analyze the most common elements of road crashes. In this report, we will model and forecast crashes in Wake County, North Carolina. To accomplish this goal, we will use the North Carolina Wake County crashes dataset from 2015 to March 2021. We were able to observe which factors have a higher cause of crashes in Wake County. We will address the above research question by using histograms of common factors leading to crashes and Time Series Forecasting crashes in Wake County.

\newpage

# Introduction 

Car crashes are something that occurs nearly every day. On average, there are 16,438 car crashes per day in the U.S. When it comes to car accidents, several factors correlate. However, it is crucial to understand some of the most frequent causes of crashes. Analyzing the most common reasons can help notify other drivers on the road so that there can be a decrease in crashes. 

When it comes to identifying car crashes, different states in the U.S experience crashes at different rates. Road accidents do not only harm the driver of the vehicle. It also affects the passenger and the pedestrians.  Most people assume that the driver gets injured the most in a road accident. Passengers and pedestrians are still dying at high rates.

From the beginning of 2015, information on car crashes has been listed in the Wake County car accident database. This data includes the driver's age, gender, vehicle type, and many more critical car accident factors. Our database is based on what police officers have recorded at the scene of the accident. Most of the data found in this database are true, but there is a small percentage of error with some of the data. Our goal is to sort out essential factors from the database to help forecast crashes in Wake County.

In this report, we will be visualizing, modeling and forecasting car crashes in Wake County. Our research questions are 1) What are the main factors that contribute to a crash? 2) Did the law change to obtaining a driver’s license due to the COVID pandemic impact crash frequency? 3) What are the conditions that lead to worse outcomes in terms of crash injuries?4) Can we forecast the daily number of crashes in Wake County? 5) Can we predict traffic hotspots in Wake County?.We assembled a dataset from the Wake County crashes database from 2015 to March 2021. Although some of the input in the database is missing, this data will help provide some information about crashes in Wake County. This will also help us forecast car crashes after 2021. 

# Background 

When it comes to car crashes, people assume that the driver was either speeding or not paying attention. It is essential to understand that there are other factors to consider when a car accident occurs. Since everything has slowly started to open after the COVID-19 pandemic, there has been an increase in car crashes. Studying the different factors of impacts can help decrease the number of accidents.

# Data, Method, and Analysis 

## Data

For this study, we utilized public data from the Wake County crashes dataset. We combined datasets from each year so that we could forecast car crashes in Wake County. This database gets updated every month and has all the specifications for the car accident. 

```{r loading-libraries, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
library(reshape2)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dslabs)
library(ggrepel)
library(ggthemes)
library(maps)
library(scales)
library(zoo)
```

```{r read-data, message=FALSE, echo=FALSE, include=FALSE}
#Code for Cortlyns computer (Mac)
 # setwd("/Users/cmorris/Desktop/dsreu2021/rstudiodirectory/ResearchProject")
 # persons <- read.csv(file="Persons_Involved_in_Crashes.csv")
 # glimpse(persons)
 #    
 # locations <- read.csv(file="Reported_Crash_Locations.csv")
 # glimpse(locations)
 #    
 # crashes <- persons %>% left_join(locations, by="key_crash")

#Code for Kelly's computer (Windows)
#library(readr)
persons <- read_csv("~/NCAT REU/Mostafa/Data/Persons_Involved_in_Crashes.csv")

locations <- read_csv("~/NCAT REU/Mostafa/Data/Reported_Crash_Locations.csv")

crashes <- persons %>% left_join(locations, by="key_crash")

# Code for Ayan's computer (Window)
#library(readr)
# persons <- read_csv("C:/Users/ayang/Desktop/Project FIles REU A&T/REU Research Project1/Persons_Involved_in_Crashes.csv")
# 
# locations <- read_csv("C:/Users/ayang/Desktop/Project FIles REU A&T/REU Research Project1/Reported_Crash_Locations.csv")
# 
# crashes <- persons %>% left_join(locations, by="key_crash")
```

## Method 

### Exploratory Data Analysis 

Exploratory data analysis was the first step in the process of understanding and finding patterns within the crash data set. We conducted univariate and bivariate explorations. 

#### Univariate Explorations 

Below are the most important univariate visualizations that we discovered. 

```{r driver-passenger-age, echo=FALSE}
crashes %>% 
  filter(PersonType == "Driver" | PersonType == "Passenger", Age != "", Age != "NA") %>% 
  ggplot() +
  geom_histogram(aes(x=Age), binwidth = 5, col="red", fill="darkgrey") +
  labs(x="Age", y="Frequency", title="Drivers' vs. Passengers' Age") +
  theme(legend.position = "top") +
  facet_wrap(~PersonType, dir = "v")
```

For the graph above we created a histogram to compare the drivers' and passengers' age. This helped us understand what age is common for drivers to get injured in a car accident. The range of drivers' ages is from about 16 to 100. The range of passengers' ages is from about 0 to 100. The range of drivers' ages is slightly larger due to the age restrictions for getting a license.

```{r crash-year, echo=FALSE}
crashes %>% 
  filter(Crash_Date_Year != "Unknown", Crash_Date_Year != "2011") %>%
  ggplot(aes(Crash_Date_Year, fill=as.factor(Crash_Date_Year))) +
  geom_bar(show.legend = FALSE) +
  geom_text(stat="count", aes(x=Crash_Date_Year, label=..count..), vjust=-0.25) + 
  geom_abline(intercept = 387995/(77/12), slope = 0, lty="dashed") +
  scale_x_continuous(breaks=2015:2021) +
  labs(title = "Frequency of Crashes by Year", 
       x = "Year", 
       y = "Count")
```

For this bar plot, we had data for the years 2015 to March 2021. In the visual above you can see that there was a high rate of car accidents from the beginning of 2015 to the end of 2019.We included a horizontal dashed line to indicate where the yearly average lies and we can see that only the years 2020 and 2021 fall below this yearly average. The year with the most crashes in 2019 but not by many. The full year with the fewest crashes in 2020 which was to be expected due to the lockdowns that took place to try and mitigate the spread of Covid-19. 2021 also looks to be on track to be another low crash year but we do not have a perfect explanation for that seeing as things are beginning to open and "return to normal" again.

```{r weather-condition, echo=FALSE}
crashes %>% 
  filter(WeatherCondition1 != "NA") %>%
  count(WeatherCondition1) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         WeatherCondition1 = reorder(WeatherCondition1,logtrans)) %>% 
  ggplot(aes(x=WeatherCondition1, y=logtrans, fill=WeatherCondition1)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Weather Condition",
       x = "Weather Condition",
       y = "Count (log10 Scale)")
```

For this visual, we also created a bar plot to determine which weather condition tends to be the most common for car crashes. After constructing the bar plot you can recognize that most crashes tend to happen when the weather is clear. This is not very surprising because in North Carolina, the majority of the time the weather is clear. An interesting note is that snow still has a relatively high frequency despite it snowing minimally in North Carolina. 

```{r vehicle-type, echo=FALSE}
crashes %>% 
  filter(VehicleType != "Unknown", VehicleType != "") %>%
  count(VehicleType) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         VehicleType = reorder(VehicleType,logtrans)) %>% 
  ggplot(aes(x=VehicleType, y=logtrans, fill=VehicleType)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) + 
  labs(title = "Frequency of Crashes by Vehicle Type", 
       x = "Vehicle Type", 
       y = "Count (log10 Scale)")
```

For the type of vehicle in a crash we had data for the years 2015 to March 2021. This bar plot helps us understand the most common vehicles and the least common vehicles in an accident in Wake County. The highest frequency of crashes occur in a passenger car. However, this is not very surprising because passenger cars are relatively common. An interesting note is that sport utility vehicles have the second highest frequency of crashes, however, sport utility vehicles are not nearly as common of a car.

#### Bivariate Explorations

...

Below are some of the most important bivariate explorations. 

...

### Identifying Hot Spots of Crashes

This section is dedicatied to identifying hot spots of crashes in Wake County, North Carolina. 

#### Word Clouds 

Below are two word clouds created to visualize the frequency of crashes on streets in Wake County, North Carolina. 

```{r tool, message=FALSE, echo=FALSE}
devtools::install_github("gaospecial/wordcloud2")
```

```{r load-library-word-cloud, message=FALSE, echo=FALSE}
#install.packages('wordcloud2')
library(wordcloud2)
```

```{r location-on-word-cloud, echo=FALSE}
location_road_name_on_freq <- crashes %>%
  count(LocationRoadNameOn)

set.seed(101)
location_road_name_on_freq %>%
  wordcloud2(shape = 'circle', backgroundColor = "black", minSize = 5) 
```

From the word cloud, we can see many crashes occur on 440, 40, and Capital Blvd. This makes sense because 440 and 40 are major interstates that connect to several major junctions and Capital Blvd is a main road in North Carolina as well. 

```{r location-at-word-cloud, echo=FALSE}
location_road_name_at_freq <- crashes %>%
  count(LocationRoadNameAt)

set.seed(102)
location_road_name_at_freq %>%
  wordcloud2(shape = 'circle', backgroundColor = "black", minSize = 5)
```

From the word cloud, we can see that 440, Capital Blvd, 40, and Blue Ridge Rd frequently have crashes that are near by. This would make sense because most of these streets are highly trafficked and populated.

#### Interactive Heat Maps 

Below are two interactive heat maps dedicated to organizing the data in a spatial manner. 

```{r load-library-interactive-map, message=FALSE, echo=FALSE}
library(ggplot2)
#install.packages("ggmap")
library(ggmap)
library(RColorBrewer)
library(plotly)
```

```{r nc-interactive-map, message=FALSE, warning=FALSE, echo=FALSE}
map_bounds <- c(-78.98, 35.51, -78.24, 36.06)

crashes_filter <- crashes %>%
  filter(LocationLatitude != "0", LocationLongitude != "0",
         LocationLatitude != "NA", LocationLatitude != "",
         LocationLongitude != "NA", LocationLongitude != "",
         LocationLatitude > 35.4, LocationLatitude < 36.2,
         LocationLongitude < -78, LocationLongitude > -79)

coords.map.nc <- get_stamenmap(map_bounds, zoom = 10, maptype = "toner-lite")

coords.map.nc <- ggmap(coords.map.nc, extent="panel", legend="none")
coords.map.nc <- coords.map.nc + stat_density2d(data=crashes_filter,  
                                          aes(x=LocationLongitude, 
                                              y=LocationLatitude, 
                                              fill=..level..),
                                          alpha = 0.3, 
                                          geom="polygon")
coords.map.nc <- coords.map.nc +   scale_fill_gradientn(colours=rev(brewer.pal(7, "RdBu")))

coords.map.nc <- coords.map.nc + theme_bw() + 
  ggtitle("Interactive Heat Map of Crashes in North Carolina (Zoomed-Out)") + 
  xlab("Longitude") + ylab("Latitude")

ggplotly(coords.map.nc) %>%
  highlight(
    "plotly_hover",
    selected = attrs_selected(line = list(color = "black"))
  )
```

In the visual above we created a heat map of common places where road accidents occur in North Carolina. 

```{r wake-interactive-map, message=FALSE, warning=FALSE, echo=FALSE}
crashes_filter <- crashes %>%
  filter(LocationLatitude != "0", LocationLongitude != "0",
         LocationLatitude != "NA", LocationLatitude != "",
         LocationLongitude != "NA", LocationLongitude != "",
         LocationLatitude > 35.4, LocationLatitude < 36.2,
         LocationLongitude < -78, LocationLongitude > -79)

map_bounds <- c(-78.8, 35.72, -78.5, 35.9) #coordinates of wake county

coords.map.wake <- get_stamenmap(map_bounds, zoom = 13, maptype = "toner-lite")

coords.map.wake <- ggmap(coords.map.wake, extent="panel")
coords.map.wake <- coords.map.wake + stat_density2d(data=crashes_filter,  
                                          aes(x=LocationLongitude, 
                                              y=LocationLatitude, 
                                              fill=..level..),
                                          alpha=0.3, 
                                          geom="polygon")

coords.map.wake <- coords.map.wake + scale_fill_gradientn(colours=rev(brewer.pal(7, "RdYlGn")))

coords.map.wake <- coords.map.wake + theme_bw() + 
  ggtitle("Heat Map of Crash Locations") + 
  xlab("Latitude") + ylab("Longitude")

ggplotly(coords.map.wake) %>%
  highlight(
    "plotly_hover",
    selected = attrs_selected(line = list(color = "black"))
  )
```

Above we created an interactive heat map for Wake County. This heat map shows popular streets where it is common to have an accident. From the Wake County heat map, we can see that crashes are much more likely to occur in the center of the city. We see that as more streets populate an area, more crashes occur. 

### Time Series Visualization 

### Time Series Forecasting 

#### Stationary Transformation 

### Statistical Analyses 

### Predictive Models 

# Results 

## Time Series Analysis Results 

### Time Series Forecasting 

### Forecasting with Covariates 

#### Daily with Pandemic Data 

#### Daily without Pandemic Data 

#### Monthly with Pandemic Data

#### Monthly without Pandemic Data 

## Predictive Modeling Results 

### Under Sampling 

### Over Sampling 

### Combination 

# Conclusion 

# Discussion 

# References

# Acknowledgements 

The authors acknowledge financial support from the NSF HBCU-UP ACE: NSF HRD-1719498 and Research Experiences for Undergraduate Students in Data Science Analytics, National Security Agency H98230-18-1-0097. 

# Appendix 

