---
title: "Visualizing, Modeling and Forecasting Crashes in Wake County, North Carolina"
author: "Cortlyn Morris, Kelly Wentzlof, Ayan Gulzar"
date: "6/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Here we will be loading necessary packages.

```{r, message=FALSE, warning=FALSE}
library(reshape2)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dslabs)
library(ggrepel)
library(ggthemes)
library(maps)
library(scales)
```

### Setting working directory.

```{r}
#setwd("/Users/cmorris/Desktop/dsreu2021/rstudiodirectory/ResearchProject")
```

### Reading and merging the data.
This code is for a Mac, if using Windows, read.csv becomes read_csv
```{r}
#Code for Cortlyn's computer
persons <- read.csv(file="Persons_Involved_in_Crashes.csv")
glimpse(persons)

locations <- read.csv(file="Reported_Crash_Locations.csv")
glimpse(locations)

crashes <- persons %>% left_join(locations, by="key_crash")

#Code for Kelly's computer 
# library(readr)
# persons <- read_csv("~/NCAT REU/Mostafa/Data/Persons_Involved_in_Crashes.csv")
# 
# locations <- read_csv("~/NCAT REU/Mostafa/Data/Reported_Crash_Locations.csv")
# glimpse(locations)
# 
# crashes <- persons %>% left_join(locations, by="key_crash")
```

## Exploratory data analysis 

### Frequency of crashes by vehicle type

The highest frequency of crashes occur in a passenger car. However, this is not very surprising because passenger cars are relatively common. An interesting note is that sport utility vehicles have the second highest frequency of crashes, however, sport utility vehicles are not nearly as common of a car.

```{r}
crashes %>% 
  filter(VehicleType != "Unknown", VehicleType != "") %>%
  count(VehicleType) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         VehicleType = reorder(VehicleType,logtrans)) %>% 
  ggplot(aes(x=VehicleType, y=logtrans, fill=VehicleType)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) + 
  labs(title = "Frequency of Crashes by Vehicle Type", 
       x = "Vehicle Type", 
       y = "Count (log10 Scale)")
```

### Frequency of crashes by weather condition.

The highest frequency of crashes occur when the weather conditions are clear. This is not very surprising because in North Carolina, the majority of the time the weather is clear. An interesting note is that snow still has a relatively high frequency despite it snowing minimally in North Carolina. 

```{r}
crashes %>% 
  filter(WeatherCondition1 != "NA") %>%
  count(WeatherCondition1) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         WeatherCondition1 = reorder(WeatherCondition1,logtrans)) %>% 
  ggplot(aes(x=WeatherCondition1, y=logtrans, fill=WeatherCondition1)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Weather Condition",
       x = "Weather Condition",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Weather Condition and if Weather Contributed to Crash

```{r}
crashes %>% 
  filter(WeatherCondition1 != "NA", WeatherContributedToCrash == "Yes") %>%
  count(WeatherCondition1) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         WeatherCondition1 = reorder(WeatherCondition1,logtrans)) %>% 
  ggplot(aes(x=WeatherCondition1, y=logtrans, fill=WeatherCondition1)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = 
  "Frequency of Crashes by Weather Condition and if Weather Contributed to Crash",
       x = "Weather Condition",
       y = "Count (log10 Scale)")
```

### Frequency of crashes by traffic control type.

The highest frequency of crashes occur when there is no traffic control present. Without a control present, drivers may be less confident in their actions. 

```{r}
crashes %>% 
  filter(TrafficControlType != "NA", TrafficControlType != "", 
         TrafficControlType != "NaN") %>%
  count(TrafficControlType) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         TrafficControlType = reorder(TrafficControlType,logtrans)) %>% 
  ggplot(aes(x=TrafficControlType, y=logtrans, fill=TrafficControlType)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Traffic Control Type",
       x = "Traffic Control Type",
       y = "Count (log10 Scale)")
```
 
### Frequency of crashes by road feature.

The highest frequency of crashes occur when there is no special road feature present. Without a road feature present, drivers may be less confident in their actions. 

```{r}
crashes %>% 
  filter(RoadFeature != "NA", RoadFeature != "", RoadFeature != "NaN") %>%
  count(RoadFeature) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         RoadFeature = reorder(RoadFeature,logtrans)) %>% 
  ggplot(aes(x=RoadFeature, y=logtrans, fill=RoadFeature)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Road Feature",
       x = "Road Feature",
       y = "Count (log10 Sclae)")
```

### Histogram for Drivers' Ages 

```{r}
crashes %>%
  filter(PersonType == "Driver", Age != "NA", Age != "NaN", Age != "") %>%
  ggplot() +
  geom_histogram(aes(x=Age), binwidth = 5, col="red", fill="darkgrey") +
  labs(title="Histogram for Drivers' Ages", x="Age", y="Frequency")
```

### Frequency of Crashes by Alcohol Result Type (including Unknown and Pending results)

```{r}
crashes %>% 
  filter(AlcoholResultType != "NA", AlcoholResultType != "", 
         AlcoholResultType != "NaN", AlcoholResultType != "Contaminated sample/unusable") %>%
  count(AlcoholResultType) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         AlcoholResultType = reorder(AlcoholResultType,logtrans)) %>% 
  ggplot(aes(x=AlcoholResultType, y=logtrans, fill=AlcoholResultType)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Alcohol Result Type",
       x = "Alcohol Result Type",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Alcohol Result Type (excluding Unknown and Pending results)

```{r}
crashes %>% 
  filter(AlcoholResultType != "NA", AlcoholResultType != "", 
         AlcoholResultType != "NaN", AlcoholResultType != 
           "Contaminated sample/unusable", AlcoholResultType != "Pending",
         AlcoholResultType != "Unknown") %>%
  count(AlcoholResultType) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         AlcoholResultType = reorder(AlcoholResultType,logtrans)) %>% 
  ggplot(aes(x=AlcoholResultType, y=logtrans, fill=AlcoholResultType)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Alcohol Result Type",
       x = "Alcohol Result Type",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Day of Week

```{r}
crashes %>% 
  filter(Crash_Date_DOW != "NA") %>%
  mutate(Crash_Date_DOW = factor(Crash_Date_DOW, levels = c("Sunday","Monday", "Tuesday", "Wednesday", 
                                   "Thursday", "Friday", "Saturday"))) %>%
  ggplot() +
  geom_bar(aes(Crash_Date_DOW, fill=Crash_Date_DOW), show.legend = FALSE) +
  geom_text(stat="count", aes(x=Crash_Date_DOW, label=..count..), vjust=-0.25) +
  labs(title = "Frequency of Crashes by Day of Week",
       x = "Day of Week",
       y = "Count")
```

### Frequency of Crashes by Month

```{r}
crashes %>% 
  filter(Crash_Date_Month != "NA") %>%
  mutate(Crash_Date_Month = factor(Crash_Date_Month, 
                                   levels = c("January","February",
                                              "March", "April",
                                              "May", "June", "July",
                                              "August", "September",
                                              "October", "November",
                                              "December"))) %>%
  ggplot() +
  geom_bar(aes(Crash_Date_Month, fill=Crash_Date_Month), show.legend = FALSE) +
  geom_text(stat="count", aes(x=Crash_Date_Month, label=..count..), vjust=-0.25) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Frequency of Crashes by Month",
       x = "Month",
       y = "Count")
```

### Frequency of Crashes by Hour

```{r}
crashes %>%
  filter(Crash_Date_Hour != "NA", Crash_Date_Hour != "") %>%
  ggplot() +
  geom_bar(aes(Crash_Date_Hour)) +
  labs(title="Frequency of Crashes by Hour", x="Hour", y="Count")
```

### Frequency of Crashes by Shift

```{r}
crashes %>%
  filter(Crash_Date_Hour != "NA", Crash_Date_Hour != "") %>%
  mutate(shift = ifelse(
    Crash_Date_Hour >= 6 & Crash_Date_Hour < 14, "Shift 1", 
    ifelse(Crash_Date_Hour >= 14 & Crash_Date_Hour < 22, "Shift 2", "Shift 3"))) %>%
  ggplot() +
  geom_bar(aes(x=shift, fill = shift)) +
  labs(title="Frequency of Crashes by Shift", x="Shift", y="Count")
```


### Frequency of Crashes by Drivers' v. Passengers' Age

The range of drivers' ages is from about 16 to 100. The range of passengers' ages is from about 0 to 100. The range of drivers' ages is slightly larger due to the age restrictions for getting a license. 

```{r}
crashes %>% 
  filter(PersonType == "Driver" | PersonType == "Passenger", Age != "", Age != "NA") %>% 
  ggplot() +
  geom_histogram(aes(x=Age), binwidth = 5, col="red", fill="darkgrey") +
  labs(x="Age", y="Frequency", title="Drivers' vs. Passengers' Age") +
  theme(legend.position = "top") +
  facet_wrap(~PersonType)
```

### Frequency of Crashes by Drivers' v. Passengers' Gender
```{r}
crashes %>% 
  filter(PersonType == "Driver" | PersonType == "Passenger", 
         Gender != "NA", Gender != "Unknown") %>% 
  ggplot() +
  geom_bar(aes(x=Gender), col="red", fill="darkgrey") +
  labs(x="Gender", y="Frequency", title="Drivers' vs. Passengers' Gender") +
  theme(legend.position = "top") +
  facet_wrap(~PersonType)
```

### Frequency of Crashes by Vehicle Seizure

About half as many vehicles were seized compared to vehicles not seized. 

```{r} 
crashes %>% 
  filter(VehicleSeizure != "NA") %>%
  count(VehicleSeizure) %>% 
  mutate(logtrans = round(log10(n), digits = 2)) %>% 
  ggplot(aes(x=VehicleSeizure, y=logtrans, fill=VehicleSeizure)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  labs(title = "Frequency of Crashes by Vehicle Seizure",
       x = "Vehicle Seizure",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Airbag Deployed

The highest frequency of crashes occured where the airbags were not deployed. It is interesting to note that the second highest frequency of crashes occured where the car(s) did not have airbags to begin with. 
Airbags did not deploy during most crashes, therefore many accidents logged were minor. 

```{r}
crashes %>% 
  filter(AirbagDeployed != "NA", AirbagDeployed != "Unknown") %>%
  count(AirbagDeployed) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         AirbagDeployed = reorder(AirbagDeployed,logtrans)) %>% 
  ggplot(aes(x=AirbagDeployed, y=logtrans, fill=AirbagDeployed)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Airbag Deployed",
       x = "Airbag Deployed",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Ejection

Most car crashes resulted in drivers and passengers not being ejected. More people were totally ejected from the vehicle than partially ejected. 

```{r}
crashes %>% 
  filter(Ejection != "NA", Ejection != "Unknown") %>%
  count(Ejection) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Ejection = reorder(Ejection,logtrans)) %>% 
  ggplot(aes(x=Ejection, y=logtrans, fill=Ejection)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Ejection",
       x = "Ejection",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Gender

There does not seem to be a significant difference between frequency of crashes between males and females. 

```{r}
crashes %>% 
  filter(Gender != "NA", Gender != "Unknown", Gender != "") %>%
  count(Gender) %>% 
  mutate(Percent = round(n/sum(n)*100,2), 
         Gender = reorder(Gender,Percent)) %>% 
  ggplot(aes(x=Gender, y=Percent, fill=Gender)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label= Percent),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Gender",
       x = "Gender",
       y = "Percent")
```

### Frequency of Crashes by Race
According to census.gov...

In the United States, 76.3 % of people are White, 13.4% of people are Black or African American, 18.5% are Hispanic or Latina, 0.2 are Native Hawaiian and Other Pacific Islander, 2.8% are two or more races, 5.9% are Asian, 1.3% are American Indian and Alaska Native, . 

In North Carolina, 70.6 % of people are White, 22.2% of people are Black or African American, 9.8% are Hispanic or Latina, 0.1 are Native Hawaiian and Other Pacific Islander, 2.3% are two or more races, 3.2% are Asian, and 1.6% are American Indian and Alaska Native.

In Wake County, North Carolina, 67.9 % of people are White, 21.0% of people are Black or African American, 10.4% are Hispanic or Latina, 0.1% are Native Hawaiian and Other Pacific Islander, 2.6% are two or more races, 7.7% are Asian, and 0.8% are American Indian and Alaska Native.

```{r}
crashes %>% 
  filter(Race != "NA", Race != "Unknown", Race != "") %>%
  count(Race) %>% 
  mutate(Percent = round(n/sum(n)*100,2), 
         Race = reorder(Race,Percent)) %>% 
  ggplot(aes(x=Race, y=Percent, fill=Race)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Percent), nudge_y = 0.05) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Race",
       x = "Race",
       y = "Percent")
```

### Frequency of Crashes by Injury

The highest frequency of crashes did not result in any injury. Following the highest frequency, the more serious the injury, the less crashes occured. 

```{r}
crashes %>% 
  filter(Injury != "NA", Injury != "Unknown") %>%
  count(Injury) %>% 
  mutate(Percent = round(n/sum(n)*100, digits = 2), 
         Injury = reorder(Injury,Percent)) %>% 
  ggplot(aes(x=Injury, y=Percent, fill=Injury)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Percent),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Injury",
       x = "Injury",
       y = "Percent")
```

### Frequency of Injuries by Person Type (old code)

```{r}
# Finding the counts of each injury by person type
crashes %>% 
  filter(Injury != "NA", Injury != "Unknown", Injury != "", 
         PersonType == "Driver") %>%
  count(Injury)

crashes %>% 
  filter(Injury != "NA", Injury != "Unknown", Injury != "", 
         PersonType == "Passenger") %>%
  count(Injury)

crashes %>% 
  filter(Injury != "NA", Injury != "Unknown", Injury != "", 
         PersonType == "Pedestrian") %>%
  count(Injury)

crashes %>% 
  filter(Injury != "NA", Injury != "Unknown", Injury != "", 
         PersonType == "Pedalcyclist") %>%
  count(Injury)

crashes %>% 
  filter(Injury != "NA", Injury != "Unknown", Injury != "", 
         PersonType == "Other*") %>%
  count(Injury)

#Creating a dataset using injury type, person type, and counts
persontype2 <- c(rep("Driver" , 5) , rep("Passenger" , 5) , rep("Pedestrian" , 5) ,
                 rep("Pedalcyclist" , 5) , rep("Other" , 5) )

injury2 <- rep(c("Type A" , "Type B" , "Type C", "Killed", "No Injury") , 5)

value <- c(851, 7295, 22020, 105, 227811, 232, 2307, 9163, 31, 72373, 183, 596, 554,
           62, 154, 19, 168, 123, 3, 68, 10, 48, 34, 1, 141)

data <- data.frame(persontype2,injury2,value)
```

```{r}
#Bar plot
data %>%
  mutate(
         injury2 = reorder(injury2, -value),
         persontype2 = factor(persontype2,
                              levels = c("Driver", "Passenger", "Pedestrian", 
                                         "Pedalcyclist", "Other"))) %>%
  ggplot(aes(fill=injury2, y=value, x=persontype2)) +
  geom_bar(position="dodge", stat="identity") +
  geom_text(aes(label=value), nudge_x = -0.35) +
  coord_flip()
```


### Frequency of Crashes by Injuries and Person Type 

```{r}
crashes %>%
  filter(Injury != "NA", Injury != "Unknown", PersonType != "NA", 
        PersonType != "NaN") %>%
  group_by(Injury) %>%
  ggplot(aes(fill=Injury, x=PersonType)) + 
  geom_bar(position="fill", stat="count") + 
  coord_flip() + 
  labs(title = "Crash Frequency by Injury and Person Type",
       x = "Person Type",
       y = "Count")
```

### Frequency of by Injury and Traffic Control Type

```{r}
crashes %>%
  filter(Injury != "NA", Injury != "Unknown", TrafficControlType != "NA", 
         TrafficControlType != "NaN") %>%
  group_by(Injury) %>%
  ggplot(aes(fill=Injury, x=TrafficControlType)) + 
  geom_bar(position="fill", stat="count") + 
  coord_flip() +
  labs(title = "Crash Frequency by Injury and Traffic Control Type",
       x = "Traffic Control Type",
       y = "Count")
```

### Frequency of Crashes by Injury and Vehicle Type

```{r}
crashes %>%
  filter(Injury != "NA", Injury != "Unknown", VehicleType != "NA", 
         VehicleType != "NaN") %>%
  group_by(Injury) %>%
  ggplot(aes(fill=Injury, x=VehicleType)) + 
  geom_bar(position="fill", stat="count") + 
  coord_flip() + 
  labs(title = "Crash Frequency by Injury and Vehicle Type",
       x = "Vehicle Type",
       y = "Count")
```

### Frequency of Crashes by Injury and Protection

```{r}
crashes %>%
  filter(Injury != "NA", Injury != "Unknown", Protection != "NA", 
         Protection != "NaN") %>%
  group_by(Injury) %>%
  ggplot(aes(fill=Injury, x=Protection)) + 
  geom_bar(position="fill", stat="count") + 
  coord_flip() + 
  labs(title = "Crash Frequency by Injury and Protection",
       x = "Protection",
       y = "Count")
```

### Frequency of Crashes by Injury and Vision Obstruction 

```{r}
crashes %>%
  filter(Injury != "NA", Injury != "Unknown", VisionObstruction != "NA", 
         VisionObstruction != "NaN") %>%
  group_by(Injury) %>%
  ggplot(aes(fill=Injury, x=VisionObstruction)) + 
  geom_bar(position="fill", stat="count") + 
  coord_flip() + 
  labs(title = "Crash Frequency by Injury and Vision Obstruction",
       x = "Vision Obstruction",
       y = "Count")
```

### Frequency of Crashes by Injury and First Harmful Event

```{r}
crashes %>%
  filter(Injury != "NA", Injury != "Unknown", FirstHarmfulEvent != "NA", 
         FirstHarmfulEvent != "NaN", FirstHarmfulEvent != "Unknown") %>%
  group_by(Injury) %>%
  ggplot(aes(fill=Injury, x=FirstHarmfulEvent)) + 
  geom_bar(position="fill", stat="count") + 
  coord_flip() + 
  labs(title = "Crash Frequency by Injury and First Harmful Event",
       x = "First Harmful Event",
       y = "Count")
```

### Frequency of Crashes by Injury and Most Harmful Event 

```{r}
crashes %>%
  filter(Injury != "NA", Injury != "Unknown", MostHarmfulEvent != "NA", 
         MostHarmfulEvent != "NaN", MostHarmfulEvent != "Unknown") %>%
  group_by(Injury) %>%
  ggplot(aes(fill=Injury, x=MostHarmfulEvent)) + 
  geom_bar(position="fill", stat="count") + 
  coord_flip() + 
  labs(title = "Crash Frequency by Injury and Most Harmful Event",
       x = "Most Harmful Event",
       y = "Count")
```

### Frequency of Crashes by Protection

The highest frequency of crashes occured when drivers and passengers wore a shoulder and lap belt protection which are common protections in more recent vehicles. However, the second highest frequency of crashes occured when drivers and passengers had used no protection. Could this be because the vehicles they were driving didn't provide protection or that the vehicles had protection, the people just chose not to use any. 

```{r}
crashes %>% 
  filter(Protection != "NA", Protection != "Unable to determine") %>%
  count(Protection) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Protection = reorder(Protection,logtrans)) %>% 
  ggplot(aes(x=Protection, y=logtrans, fill=Protection)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Protection",
       x = "Protection",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Trapped

More crashes resulted in people in the vehicle not being trapped than people in the vehicle being trapped. 

```{r}
crashes %>% 
  filter(Trapped != "NA", Trapped != "Unknown") %>%
  count(Trapped) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Trapped = reorder(Trapped,logtrans)) %>% 
  ggplot(aes(x=Trapped, y=logtrans, fill=Trapped)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Trapped",
       x = "Trapped",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Vision Obstruction

The highest frequency of crashes occured when there was was no vision obstruction. 

```{r}
crashes %>% 
  filter(VisionObstruction != "NA", VisionObstruction != "Unknown") %>%
  count(VisionObstruction) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         VisionObstruction = reorder(VisionObstruction,logtrans)) %>% 
  ggplot(aes(x=VisionObstruction, y=logtrans, fill=VisionObstruction)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Vision Obstruction",
       x = "Vision Obstruction",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Contributing Circumstance 1

The highest frequency of crashes had no contributing circumstances indicated. 

```{r}
crashes %>% 
  filter(ContributingCircumstance1 != "NA", ContributingCircumstance1 != "Unknown") %>%
  count(ContributingCircumstance1) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         ContributingCircumstance1 = reorder(ContributingCircumstance1,logtrans)) %>% 
  ggplot(aes(x=ContributingCircumstance1, y=logtrans, fill=ContributingCircumstance1)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Contributing Circumstance 1",
       x = "Contributing Circumstance 1",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Contributing Circumstance 2

The highest frequency of crashes had inattention as a contributing circumstance 2. 

```{r}
crashes %>% 
  filter(ContributingCircumstance2 != "NA", 
         ContributingCircumstance2 != "Unknown") %>%
  count(ContributingCircumstance2) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         ContributingCircumstance2 = reorder(ContributingCircumstance2,
                                             logtrans)) %>% 
  ggplot(aes(x=ContributingCircumstance2, y=logtrans, 
             fill=ContributingCircumstance2)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Contributing Circumstance 2",
       x = "Contributing Circumstance 2",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Contributing Circumstance 3

The higest frequency of crashes had inattention as a contributing circumstance 3. 

```{r}
crashes %>% 
  filter(ContributingCircumstance3 != "NA", 
         ContributingCircumstance3 != "Unknown") %>%
  count(ContributingCircumstance3) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         ContributingCircumstance3 = 
           reorder(ContributingCircumstance3,logtrans)) %>% 
  ggplot(aes(x=ContributingCircumstance3, y=logtrans, 
             fill=ContributingCircumstance3)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Contributing Circumstance 3",
       x = "Contributing Circumstance 3",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Location Relation to Road 

```{r}
crashes %>% 
  filter(LocationRelationToRoad != "NA", LocationRelationToRoad != "Unknown") %>%
  count(LocationRelationToRoad) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         LocationRelationToRoad = reorder(LocationRelationToRoad,logtrans)) %>% 
  ggplot(aes(x=LocationRelationToRoad, y=logtrans, fill=LocationRelationToRoad)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Location of Crash in Relation to Road",
       x = "Location Relation to Road",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Location In Near Indicator

```{r}
crashes %>% 
  filter(LocationInNearIndicator != "NA", LocationInNearIndicator != "Unknown") %>%
  count(LocationInNearIndicator) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         LocationInNearIndicator = reorder(LocationInNearIndicator,logtrans)) %>% 
  ggplot(aes(x=LocationInNearIndicator, y=logtrans, fill=LocationInNearIndicator)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Crashes Near Indicator",
       x = "Location in Near Indicator",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Location Ramp Indicator 

```{r}
crashes %>% 
  filter(LocationRampIndicator != "NA", LocationRampIndicator != "Unknown") %>%
  count(LocationRampIndicator) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         LocationRampIndicator = reorder(LocationRampIndicator,logtrans)) %>% 
  ggplot(aes(x=LocationRampIndicator, y=logtrans, fill=LocationRampIndicator)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Crash Location from Ramp",
       x = "Location Ramp Indicator",
       y = "Count (log10 Scale)")
```

### Histogram of Crash Location Feet From Road 

```{r}
crashes %>%
  filter(LocationFeetFromRoad != "NA", LocationFeetFromRoad != "Unknown", 
         LocationFeetFromRoad != "") %>%
  ggplot() +
  geom_histogram(aes(x=LocationFeetFromRoad), binwidth = 100, col="red", fill="darkgrey") +
  labs(title="Histogram for Location Feet From Road")
```

### Frequency of Crashes by Location Direction From Road

```{r}
crashes %>% 
  filter(LocationDirectionFromRoad != "NA", LocationDirectionFromRoad != "Unknown",
         LocationDirectionFromRoad != "") %>%
  count(LocationDirectionFromRoad) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         LocationDirectionFromRoad= reorder(LocationDirectionFromRoad,logtrans)) %>% 
  ggplot(aes(x=LocationDirectionFromRoad, y=logtrans, fill=LocationDirectionFromRoad)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Direction of Crash",
       x = "Location Direction From Road",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Location From Indicator

```{r}
crashes %>% 
  filter(LocationAtFromIndicator != "NA", LocationAtFromIndicator != "Unknown") %>%
  count(LocationAtFromIndicator) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         LocationAtFromIndicator= reorder(LocationAtFromIndicator,logtrans)) %>% 
  ggplot(aes(x=LocationAtFromIndicator, y=logtrans, fill=LocationAtFromIndicator)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Indicator Crash Location",
       x = "Location At/From Indicator",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Location Direction to Road

```{r}
crashes %>% 
  filter(LocationDirectionToRoad != "NA", LocationDirectionToRoad != "Unknown") %>%
  count(LocationDirectionToRoad) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         LocationDirectionToRoad= reorder(LocationDirectionToRoad,logtrans)) %>% 
  ggplot(aes(x=LocationDirectionToRoad, y=logtrans, fill=LocationDirectionToRoad)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Direction of Crash From Road",
       x = "Location Direction To Road",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by First Harmful Event

```{r}
crashes %>% 
  filter(FirstHarmfulEvent != "NA", FirstHarmfulEvent != "Unknown") %>%
  count(FirstHarmfulEvent) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         FirstHarmfulEvent= reorder(FirstHarmfulEvent,logtrans)) %>% 
  ggplot(aes(x=FirstHarmfulEvent, y=logtrans, fill=FirstHarmfulEvent)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "First Harmful Crash Event ",
       x = "First Harmful Event",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Most Harmful Event

```{r}
crashes %>% 
  filter(MostHarmfulEvent != "NA", MostHarmfulEvent != "Unknown") %>%
  count(MostHarmfulEvent) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         MostHarmfulEvent= reorder(MostHarmfulEvent,logtrans)) %>% 
  ggplot(aes(x=MostHarmfulEvent, y=logtrans, fill=MostHarmfulEvent)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Most Harmful Crash Event ",
       x = "Most Harmful Event",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Road Classification

```{r}
crashes %>% 
  filter(RoadClassification != "NA", RoadClassification != "Unknown", 
         RoadClassification != "") %>%
  count(RoadClassification) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         RoadClassification= reorder(RoadClassification,logtrans)) %>% 
  ggplot(aes(x=RoadClassification, y=logtrans, fill=RoadClassification)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Classification of the Road ",
       x = "Road Classification",
       y = "Count (log10 Scale)")
```

### Frequency of Crashes by Crash Date Year

```{r}
crashes %>% 
  filter(Crash_Date_Year != "Unknown", Crash_Date_Year != "2011") %>%
  ggplot(aes(Crash_Date_Year, fill=Crash_Date_Year)) +
  geom_bar(aes(Crash_Date_Year, fill = Crash_Date_Year), show.legend = FALSE) +
  geom_text(stat="count", aes(x=Crash_Date_Year, label=..count..), vjust=-0.25) + 
  geom_abline(intercept = 387995/(77/12), slope = 0, lty="dashed") +
  labs(title = "Frequency of Crashes by Year", 
       x = "Year", 
       y = "Count")
```

### Number of Drivers per Crash

```{r}
crashes %>% 
  filter(PersonType == "Driver") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(Drivers = n, Frequency = nn) %>%
  mutate(Drivers = reorder(Drivers, Frequency)) %>%
  ggplot(aes(x=Drivers, y=Frequency, fill=Drivers)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y=0.1) +
  coord_flip() +
  labs(title = "Number of Drivers per Crash",
       x = "Number of Drivers",
       y = "Count")
```

### Number of Passengers per Crash

```{r}
crashes %>% 
  filter(PersonType == "Passenger") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(Passengers = n, Frequency = nn) %>%
  mutate(Passengers = reorder(Passengers,Frequency)) %>% 
  ggplot(aes(x=Passengers, y=Frequency, fill=Passengers)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y=0.5) +
  coord_flip() +
  labs(title = "Number of Passengers per Crash",
       x = "Number of Passengers",
       y = "Count")
```

### Number of Pedestrians per Crash

```{r}
crashes %>% 
  filter(PersonType == "Pedestrian") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>% 
  rename(Pedestrians = n, Frequency = nn) %>%
  mutate(Pedestrians = reorder(Pedestrians,Frequency)) %>% 
  ggplot(aes(x=Pedestrians, y=Frequency, fill=Pedestrians)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y=0.5) +
  coord_flip() +
  labs(title = "Number of Pedestrians per Crash",
       x = "Number of Pedestrians",
       y = "Count")
```

### Number of Pedalcyclists per Crash

```{r}
crashes %>% 
  filter(PersonType == "Pedalcyclist") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(Pedalcyclists = n, Frequency = nn) %>%
  mutate(Pedalcyclists = reorder(Pedalcyclists,Frequency)) %>% 
  ggplot(aes(x=Pedalcyclists, y=Frequency, fill=Pedalcyclists)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y=0.5) +
  coord_flip() +
  labs(title = "Number of Pedalcyclists per Crash",
       x = "Number of Pedalcyclists",
       y = "Count")
```

### Number of Other People per Crash

```{r}
crashes %>% 
  filter(PersonType == "Other*") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(Other = n, Frequency = nn) %>%
  mutate(Other = reorder(Other,Frequency)) %>% 
  ggplot(aes(x=Other, y=Frequency, fill=Other)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y=0.5) +
  coord_flip() +
  labs(title = "Number of Others per Crash",
       x = "Number of Others",
       y = "Count")
```

### Number of Unknown People per Crash

```{r}
crashes %>% 
  filter(PersonType == "Unknown") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(Unknown = n, Frequency = nn) %>%
  mutate(Unknown = reorder(Unknown,Frequency)) %>% 
  ggplot(aes(x=Unknown, y=Frequency, fill=Unknown)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y=0.5) +
  coord_flip() +
  labs(title = "Number of Unknown per Crash",
       x = "Number of Unknown",
       y = "Count")
```

### Number of People Killed per Crash

```{r}
crashes %>% 
  filter(Injury == "Killed") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(Killed = n, Frequency = nn) %>%
  mutate(Killed = reorder(Killed, Frequency)) %>%
  ggplot(aes(x=Killed, y=Frequency, fill=Killed)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y = 0.5) +
  coord_flip() +
  labs(title = "Number Killed per Crash",
       x = "Number Killed",
       y = "Count")
```

### Number of Type A Injuries per Crash

```{r}
crashes %>% 
  filter(Injury == "A type injury (disabling)") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(TypeA = n, Frequency = nn) %>%
  mutate(TypeA = reorder(TypeA, Frequency)) %>%
  ggplot(aes(x=TypeA, y=Frequency, fill=TypeA)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y = 0.5) +
  coord_flip() +
  labs(title = "Number of Type A Injuries per Crash",
       x = "Number of Type A Injuries",
       y = "Count")
```

### Number of Type B Injuries per Crash 

```{r}
crashes %>% 
  filter(Injury == "B type injury (evident)") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(TypeB = n, Frequency = nn) %>%
  mutate(TypeB = reorder(TypeB, Frequency)) %>%
  ggplot(aes(x=TypeB, y=Frequency, fill=TypeB)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y = 0.5) +
  coord_flip() +
  labs(title = "Number of Type B Injuries per Crash",
       x = "Number of Type B Injuries",
       y = "Count")
```

### Number of Type C Injuries per Crash

```{r}
crashes %>% 
  filter(Injury == "C type injury (possible)") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(TypeC = n, Frequency = nn) %>%
  mutate(TypeC = reorder(TypeC, Frequency)) %>%
  ggplot(aes(x=TypeC, y=Frequency, fill=TypeC)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y = 0.5) +
  coord_flip() +
  labs(title = "Number of Type C Injuries per Crash",
       x = "Number of Type C Injuries",
       y = "Count")
```

### Number of People with No Injuries per Crash 

```{r}
crashes %>% 
  filter(Injury == "No injury") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(NoInjury = n, Frequency = nn) %>%
  mutate(NoInjury = reorder(NoInjury, Frequency)) %>%
  ggplot(aes(x=NoInjury, y=Frequency, fill=NoInjury)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y = 0.5) +
  coord_flip() +
  labs(title = "Number of People with No Injuries per Crash",
       x = "Number of People with No Injuries",
       y = "Count")
```

### Number of Unknown Injuries per Crash

```{r}
crashes %>% 
  filter(Injury == "Unknown") %>%
  group_by(key_crash) %>%
  count() %>% 
  group_by(n) %>%
  count() %>%
  rename(Unknown = n, Frequency = nn) %>%
  mutate(Unknown = reorder(Unknown, Frequency)) %>%
  ggplot(aes(x=Unknown, y=Frequency, fill=Unknown)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=Frequency),nudge_y = 0.5) +
  coord_flip() +
  labs(title = "Number of Unknown Injuries per Crash",
       x = "Number of Unknown Injuries",
       y = "Count")
```

### Map of Crash Locations

```{r}
crashes %>% 
  filter(LocationLatitude != "0", LocationLongitude != "0",
         LocationLatitude != "NA", LocationLatitude != "",
         LocationLongitude != "NA", LocationLongitude != "",
         LocationLatitude > 35.4, LocationLatitude < 36.2,
         LocationLongitude < -78, LocationLongitude > -79) %>%
  ggplot(aes(LocationLongitude, LocationLatitude)) +
  borders('county', 'north carolina', fill = "light blue") +
  geom_point(size=0.5, col = 'black', show.legend = F) +
  coord_quickmap()
```

## Visualizing text data using word clouds 

### Installing and loading necessary packages 

```{r}
devtools::install_github("gaospecial/wordcloud2")
```

```{r}
#install.packages('wordcloud2')
library(wordcloud2)
```

### Word Cloud for Location Road Name On  

```{r}
location_road_name_on_freq <- crashes %>%
  count(LocationRoadNameOn)

set.seed(101)
location_road_name_on_freq %>%
  wordcloud2(shape = 'circle', backgroundColor = "black", minSize = 5)
```

### Word cloud for location road name at

```{r}
location_road_name_at_freq <- crashes %>%
  count(LocationRoadNameAt)

set.seed(102)
location_road_name_at_freq %>%
  wordcloud2(shape = 'circle', backgroundColor = "black", minSize = 5)
```

