---
title: "Visualizing, Modeling and Forecasting Crashes in Wake County, North Carolina"
author: "Cortlyn Morris, Kelly Wentzlof, Ayan Gulzar"
date: "6/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Here we will be loading necessary packages.

```{r, message=FALSE, warning=FALSE}
library(reshape2)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dslabs)
library(ggrepel)
library(ggthemes)
```

### Setting working directory.

```{r}
setwd("/Users/cmorris/Desktop/dsreu2021/rstudiodirectory/ResearchProject")
```

### Reading and merging the data.

```{r}
persons <- read.csv(file="Persons_Involved_in_Crashes.csv")
glimpse(persons)

locations <- read.csv(file="Reported_Crash_Locations.csv")
glimpse(locations)

crashes <- persons %>% left_join(locations, by="key_crash")
```

## Exploratory data analysis using log transformed data.

### Frequency of crashes by vehicle type
#### 
```{r}
crashes %>% 
  filter(VehicleType != "Unknown", VehicleType != "") %>%
  count(VehicleType) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         VehicleType = reorder(VehicleType,logtrans)) %>% 
  ggplot(aes(x=VehicleType, y=logtrans, fill=VehicleType)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) + 
  labs(title = "Frequency of Crashes by Vehicle Type", 
       x = "Vehicle Type", 
       y = "Count (log10 Scale)")
```

### Frequency of crashes by weather condition.
#### 
```{r}
crashes %>% 
  filter(WeatherCondition1 != "NA") %>%
  count(WeatherCondition1) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         WeatherCondition1 = reorder(WeatherCondition1,logtrans)) %>% 
  ggplot(aes(x=WeatherCondition1, y=logtrans, fill=WeatherCondition1)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Weather Condition",
       x = "Weather Condition",
       y = "Count (log10 Scale)")
```

### Frequency of crashes by traffic control type.
#### 
```{r}
crashes %>% 
  filter(TrafficControlType != "NA", TrafficControlType != "", 
         TrafficControlType != "NaN") %>%
  count(TrafficControlType) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         TrafficControlType = reorder(TrafficControlType,logtrans)) %>% 
  ggplot(aes(x=TrafficControlType, y=logtrans, fill=TrafficControlType)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Traffic Control Type",
       x = "Traffic Control Type",
       y = "Count (log10 Scale)")
```
 
### Frequency of crashes by road feature.
#### 
```{r}
crashes %>% 
  filter(RoadFeature != "NA", RoadFeature != "", RoadFeature != "NaN") %>%
  count(RoadFeature) %>% 
  mutate(logtrans = round(log10(n), digits = 2),
         RoadFeature = reorder(RoadFeature,logtrans)) %>% 
  ggplot(aes(x=RoadFeature, y=logtrans, fill=RoadFeature)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() + 
  geom_text(aes(label=logtrans),nudge_y=0.5) +
  labs(title = "Frequency of Crashes by Road Feature",
       x = "Road Feature",
       y = "Count (log10 Sclae)")
```

###Frequency of Crashes by Drivers' v. Passengers' Age

```{r}
crashes %>% 
  filter(PersonType == "Driver" | PersonType == "Passenger") %>% 
  ggplot() +
  geom_histogram(aes(x=Age), binwidth = 5, col="red", fill="darkgrey") +
  labs(x="Age", y="Frequency", title="Drivers' vs. Passengers' Age") +
  theme(legend.position = "top") +
  facet_wrap(~PersonType)
```

###Frequency of Crashes by Vehicle Seizure

```{r} 
crashes %>% 
  filter(VehicleSeizure != "NA") %>%
  count(VehicleSeizure) %>% 
  mutate(logtrans = round(log10(n), digits = 2)) %>% 
  ggplot(aes(x=VehicleSeizure, y=logtrans, fill=VehicleSeizure)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  labs(title = "Frequency of Crashes by Vehicle Seizure",
       x = "Vehicle Seizure",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Airbag Deployed

```{r}
crashes %>% 
  filter(AirbagDeployed != "NA", AirbagDeployed != "Unknown") %>%
  count(AirbagDeployed) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         AirbagDeployed = reorder(AirbagDeployed,logtrans)) %>% 
  ggplot(aes(x=AirbagDeployed, y=logtrans, fill=AirbagDeployed)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Airbag Deployed",
       x = "Airbag Deployed",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Ejection

```{r}
crashes %>% 
  filter(Ejection != "NA", Ejection != "Unknown") %>%
  count(Ejection) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Ejection = reorder(Ejection,logtrans)) %>% 
  ggplot(aes(x=Ejection, y=logtrans, fill=Ejection)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Ejection",
       x = "Ejection",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Gender

```{r}
crashes %>% 
  filter(Gender != "NA", Gender != "Unknown") %>%
  count(Gender) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Gender = reorder(Gender,logtrans)) %>% 
  ggplot(aes(x=Gender, y=logtrans, fill=Gender)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Gender",
       x = "Gender",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Race

```{r}
crashes %>% 
  filter(Race != "NA", Race != "Unknown") %>%
  count(Race) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Race = reorder(Race,logtrans)) %>% 
  ggplot(aes(x=Race, y=logtrans, fill=Race)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Race",
       x = "Race",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Injury

```{r}
crashes %>% 
  filter(Injury != "NA", Injury != "Unknown") %>%
  count(Injury) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Injury = reorder(Injury,logtrans)) %>% 
  ggplot(aes(x=Injury, y=logtrans, fill=Injury)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Injury",
       x = "Injury",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Protection

```{r}
crashes %>% 
  filter(Protection != "NA", Protection != "Unable to determine") %>%
  count(Protection) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Protection = reorder(Protection,logtrans)) %>% 
  ggplot(aes(x=Protection, y=logtrans, fill=Protection)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Protection",
       x = "Protection",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Trapped

```{r}
crashes %>% 
  filter(Trapped != "NA", Trapped != "Unknown") %>%
  count(Trapped) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         Trapped = reorder(Trapped,logtrans)) %>% 
  ggplot(aes(x=Trapped, y=logtrans, fill=Trapped)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Trapped",
       x = "Trapped",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Vision Obstruction

```{r}
crashes %>% 
  filter(VisionObstruction != "NA", VisionObstruction != "Unknown") %>%
  count(VisionObstruction) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         VisionObstruction = reorder(VisionObstruction,logtrans)) %>% 
  ggplot(aes(x=VisionObstruction, y=logtrans, fill=VisionObstruction)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Vision Obstruction",
       x = "Vision Obstruction",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Contributing Circumstance 1

```{r}
crashes %>% 
  filter(ContributingCircumstance1 != "NA", ContributingCircumstance1 != "Unknown") %>%
  count(ContributingCircumstance1) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         ContributingCircumstance1 = reorder(ContributingCircumstance1,logtrans)) %>% 
  ggplot(aes(x=ContributingCircumstance1, y=logtrans, fill=ContributingCircumstance1)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Contributing Circumstance 1",
       x = "Contributing Circumstance 1",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Contributing Circumstance 2

```{r}
crashes %>% 
  filter(ContributingCircumstance2 != "NA", 
         ContributingCircumstance2 != "Unknown") %>%
  count(ContributingCircumstance2) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         ContributingCircumstance2 = reorder(ContributingCircumstance2,
                                             logtrans)) %>% 
  ggplot(aes(x=ContributingCircumstance2, y=logtrans, 
             fill=ContributingCircumstance2)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Contributing Circumstance 2",
       x = "Contributing Circumstance 2",
       y = "Count (log10 Scale)")
```

###Frequency of Crashes by Contributing Circumstance 3

```{r}
crashes %>% 
  filter(ContributingCircumstance3 != "NA", 
         ContributingCircumstance3 != "Unknown") %>%
  count(ContributingCircumstance3) %>% 
  mutate(logtrans = round(log10(n), digits = 2), 
         ContributingCircumstance3 = 
           reorder(ContributingCircumstance3,logtrans)) %>% 
  ggplot(aes(x=ContributingCircumstance3, y=logtrans, 
             fill=ContributingCircumstance3)) +
  geom_bar(stat = "identity", show.legend = FALSE) + 
  geom_text(aes(label=logtrans),nudge_y=0.2) +
  coord_flip() +
  labs(title = "Frequency of Crashes by Contributing Circumstance 3",
       x = "Contributing Circumstance 3",
       y = "Count (log10 Scale)")
```
