---
title: "Time Series Forecast"
author: "Kelly Wentzlof"
date: "7/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading-libraries, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
library(reshape2)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(dslabs)
library(ggrepel)
library(ggthemes)
library(maps)
library(scales)
library(zoo)
library(forecast)
library(urca)
library(sf)
library(TTR)
library(fpp2)
```

```{r read-data, message=FALSE, echo=FALSE, include=FALSE}
#Code for Cortlyns computer (Mac)
 # setwd("/Users/cmorris/Desktop/dsreu2021/rstudiodirectory/ResearchProject")
 # persons <- read.csv(file="Persons_Involved_in_Crashes.csv")
 # glimpse(persons)
 #    
 # locations <- read.csv(file="Reported_Crash_Locations.csv")
 # glimpse(locations)
 #    
 # crashes <- persons %>% left_join(locations, by="key_crash")

#Code for Kelly's computer (Windows)
#library(readr)
persons <- read_csv("~/NCAT REU/Mostafa/Data/Persons_Involved_in_Crashes.csv")

locations <- read_csv("~/NCAT REU/Mostafa/Data/Reported_Crash_Locations.csv")

crashes <- persons %>% left_join(locations, by="key_crash")

# Code for Ayan's computer (Window)
#library(readr)
# persons <- read_csv("C:/Users/ayang/Desktop/Project FIles REU A&T/REU Research Project1/Persons_Involved_in_Crashes.csv")
# 
# locations <- read_csv("C:/Users/ayang/Desktop/Project FIles REU A&T/REU Research Project1/Reported_Crash_Locations.csv")
# 
# crashes <- persons %>% left_join(locations, by="key_crash")
```

```{r data-frame, echo=FALSE, message=FALSE}
#daily with pandemic 
crashes_ts = crashes %>%
  filter(as.Date(crash_date) >= "2015/01/01" & as.Date(crash_date) <= "2021/05/31") %>%
  separate(crash_date, 
           into = c("Date", "Hour"), sep = 11) %>%
  group_by(Date) %>%
  summarize(count = length(unique(key_crash)))

crashests2 <- ts(crashes_ts$count, start = c(2015,1), end = c(2021,153),
                frequency = 365)

#daily without pandemic 
crashes_ts.noncovid = crashes %>%
  filter(as.Date(crash_date) >= "2015/01/01" & as.Date(crash_date) <= "2020/03/01") %>%
  separate(crash_date, 
           into = c("Date", "Hour"), sep = 11) %>%
  group_by(Date) %>%
  summarize(count = length(unique(key_crash)))

crashests.noncovid <- ts(crashes_ts.noncovid$count, start = c(2015,1), 
                         end = c(2020,59), frequency = 365)

#monthly with pandemic 
crashes_mts = crashes %>%
  filter(as.Date(crash_date) >= "2015/01/01" & as.Date(crash_date) <= "2021/05/31") %>%
  separate(crash_date, 
           into = c("Date", "Hour"), sep = 11) %>%
  group_by(Date) %>%
  summarize(count = length(unique(key_crash))) %>%
  separate(Date, into = c("Year", "Month", "Day"), sep = "/") %>%
  group_by(Year, Month) %>%
  summarise(mcount = sum(count)) %>%
  tidyr::spread(key=Month, value=mcount)

crashes_mts = as.data.frame(crashes_mts)

rownames(crashes_mts) = seq(2015, 2021)

crashes_mts3 = crashes_mts %>%
  select(02:13)

crashes_mts2 <- ts(c(t(crashes_mts3)), frequency=12)

crashes_mts4 <- window(crashes_mts2, start=c(1,01), end=c(7,05), frequency=12)

#monthly without pandemic 
crashes_mts.noncovid = crashes %>%
  filter(as.Date(crash_date) >= "2015/01/01" & as.Date(crash_date) <= "2020/02/29") %>%
  separate(crash_date, 
           into = c("Date", "Hour"), sep = 11) %>%
  group_by(Date) %>%
  summarize(count = length(unique(key_crash))) %>%
  separate(Date, into = c("Year", "Month", "Day"), sep = "/") %>%
  group_by(Year, Month) %>%
  summarise(mcount = sum(count)) %>%
  tidyr::spread(key=Month, value=mcount)

crashes_mts.noncovid = as.data.frame(crashes_mts.noncovid)

rownames(crashes_mts.noncovid) = seq(2015, 2020)

crashes_mts3.noncovid = crashes_mts.noncovid %>%
  select(02:13)

crashes_mts2.noncovid <- ts(c(t(crashes_mts3.noncovid)), frequency=12)

crashes_mts4.noncovid <- window(crashes_mts2.noncovid, start=c(1,01), 
                                end=c(6,02), frequency=12)
```

## Time Series Forecasting 

### Daily with Pandemic Data 

#### Stationarity Check 

```{r stationarity-trans-daily-covid}
##Looking at stationarity of crashests2 daily time series 
library(urca)
acf(crashests2)
summary(ur.kpss(crashests2))

ndiffs(crashests2) #1
nsdiffs(crashests2) #0

#Attempt at differencing 
dif_crashests2 <- diff(crashests2)

#Looking at stationarity of first difference 
acf(dif_crashests2)
summary(ur.kpss(dif_crashests2))

ndiffs(dif_crashests2) #0
nsdiffs(dif_crashests2) #0
```

```{r stat-trans-daily-covid-plot}
cbind("Crashes" = crashests2,
      "First differenced" = diff(crashests2)) %>%
  autoplot(facets=TRUE) +
  xlab("Date") + ylab("") +
  ggtitle("Stationarity Transformations of Daily Pandemic Data")
```

#### TBATS Model 

```{r tbats-stationary, echo=FALSE, message=FALSE, include=FALSE}
#Stationary

#reload libraries 
crashes_ts = crashes %>%
  filter(as.Date(crash_date) >= "2015/01/01" & as.Date(crash_date) <= "2021/05/31") %>%
  separate(crash_date, 
           into = c("Date", "Hour"), sep = 11) %>%
  group_by(Date) %>%
  summarize(count = length(unique(key_crash)))

crashests2 <- ts(crashes_ts$count, start = c(2015,1), end = c(2021,153),
                frequency = 365)

dif_crashests2 <- diff(crashests2)

dif_y <- msts(dif_crashests2, seasonal.periods=c(7,365.25))

dif_fit_tbats_covid2 <- tbats(dif_y)

dif_fc_tbats_covid2 <- forecast::forecast(dif_fit_tbats_covid2, h=214)

dYhat <- dif_fc_tbats_covid2$mean #point forecast values 

#crashests2
#dif_crashests2<-diff(crashests2)
#z0<-cumsum(c(crashests2[1], dif_crashests2))
#all(crashests2==z0) #TRUE

#dif_crashests2 <- diff(crashests2) #already ran earlier
#dYhat <- dif_fc_tbats_covid2$mean #already ran earlier 
Yhat <- cumsum(c(crashests2[length(crashests2)],dYhat))
Yhat <- ts(Yhat, start = c(2021, 152), frequency=365)
```

```{r tbats-stationary-plot, echo=FALSE, message=FALSE}
#Stationary
autoplot(crashests2) + 
  autolayer(Yhat, series="Point Forecasts") + 
  ggtitle("TBATS Forecasting Model for Daily Car Crashes (With Stationary Pandemic Data)") +
  xlab("Date") + ylab("Crashes") + 
  theme(title = element_text(size = 9), legend.position = "bottom")
```

#### Multi-Step TBATS 

```{r tbats-ms-stationary, echo=FALSE, message=FALSE, include=FALSE}
#Stationary
dif_y2 <- msts(dif_crashests2, seasonal.periods=c(7,365.25))

dif_training_TBATS2 <- subset(dif_y2, end=length(dif_y2)-151)

dif_test_TBATS2 <- subset(dif_y2, start=length(dif_y2)-150)

dif_crashes.train_TBATS2 <- tbats(dif_training_TBATS2)

dif_fc_train_TBATS2 <- forecast(dif_crashes.train_TBATS2, h=151)

dYhat_ms_TBATS <- dif_fc_train_TBATS2$mean

y2 <- msts(crashests2, seasonal.periods=c(7,365.25))

training_TBATS2 <- subset(y2, end=length(y2)-151)

test_TBATS2 <- subset(y2, start=length(y2)-150)
  
#Yhat_ms_TBATS <- cumsum(c(crashests2[length(crashests2)], dYhat_ms_TBATS))
Yhat_ms_TBATS <- cumsum(c(training_TBATS2[length(training_TBATS2)], dYhat_ms_TBATS))

Yhat_ms_TBATS <- ts(Yhat_ms_TBATS, start = c(2021, 1), frequency = 365)
```

```{r tbats-ms-stationary-plot, echo=FALSE}
#Stationary 
autoplot(training_TBATS2) + 
  autolayer(Yhat_ms_TBATS, series="Point Forecasts") + 
  autolayer(test_TBATS2, series="Test Set", alpha = 0.7) + 
  ggtitle("Multi-Step TBATS Daily Forecasts of Crashes (With Stationary Pandemic Data)") + 
  xlab("Date") + ylab("Crashes") + 
  theme(legend.position = "bottom")

dif_crashes.test_TBATS2 <- tbats(dif_test_TBATS2)

accuracy(dif_crashes.test_TBATS2)
```

#### ARIMA Model 




#### Multi-Step ARIMA 


#### Combination 


### Daily without Pandemic Data 

#### Stationarity Check 

```{r stationarity-trans-daily-noncovid}
##Looking at stationarity of crashests.noncovid daily time series 
library(urca)
acf(crashests.noncovid)
summary(ur.kpss(crashests.noncovid))

ndiffs(crashests.noncovid) #1
nsdiffs(crashests.noncovid) #0

#Attempt at differencing (noncovid)
dif_crashests.noncovid <- diff(crashests.noncovid)

#Looking at stationarity of first difference (noncovid)
acf(dif_crashests.noncovid)
summary(ur.kpss(dif_crashests.noncovid))

ndiffs(dif_crashests.noncovid) #0
nsdiffs(dif_crashests.noncovid) #0
```

```{r stat-trans-daily-noncovid-plot}
cbind("Crashes" = crashests.noncovid,
      "First differenced" = diff(crashests.noncovid)) %>%
  autoplot(facets=TRUE) +
  xlab("Date") + ylab("") +
  ggtitle("Stationarity Transformation of Daily NonPandemic Data")
```

#### TBATS Model 

```{r tbats-noncovid-stationary, echo=FALSE, message=FALSE, include=FALSE}
#Stationary

crashes_ts.noncovid = crashes %>%
  filter(as.Date(crash_date) >= "2015/01/01" & as.Date(crash_date) <= "2020/03/01") %>%
  separate(crash_date, 
           into = c("Date", "Hour"), sep = 11) %>%
  group_by(Date) %>%
  summarize(count = length(unique(key_crash)))

crashests.noncovid <- ts(crashes_ts.noncovid$count, start = c(2015,1), 
                         end = c(2020,59), frequency = 365)

dif_crashests.noncovid <- diff(crashests.noncovid)

dif_y_noncovid <- msts(dif_crashests.noncovid, seasonal.periods=c(7,365.25))

dif_fit_tbats_noncovid <- tbats(dif_y_noncovid)

dif_fc_tbats_noncovid <- forecast::forecast(dif_fit_tbats_noncovid, h=671)

dYhat_noncovid <- dif_fc_tbats_noncovid$mean #point forecast values 

#crashests.noncovid
#dif_crashests.noncovid<-diff(crashests.noncovid)
#z0<-cumsum(c(crashests.noncovid[1], dif_crashests.noncovid))
#all(crashests.noncovid==z0) #TRUE

#dif_crashests2 <- diff(crashests2) #already ran earlier
#dYhat <- dif_fc_tbats_covid2$mean #already ran earlier 
Yhat_noncovid <- cumsum(c(crashests.noncovid[length(crashests.noncovid)],dYhat_noncovid))
Yhat_noncovid <- ts(Yhat, start = c(2020, 60), frequency=365)
```

```{r tbats-noncovid-stationary-plot, echo=FALSE, message=FALSE}
#Stationary

true_noncovid_crashes <- crashes_ts %>%
  filter(Date >= as.Date("2020/03/01") & Date <= as.Date("2021/05/31"))

true_noncovid_crashes <- ts(true_noncovid_crashes$count, 
                            start = c(2020,60), end = c(2021,153), 
                            frequency = 365)

autoplot(crashests.noncovid) + 
  autolayer(true_noncovid_crashes, alpha=0.7, series="True Count") +
  autolayer(Yhat_noncovid, alpha=0.6, series="Point Forecasts") + 
  ggtitle("TBATS Forecasting Model for Daily Car Crashes (Stationary Without Pandemic Data)") +
  xlab("Date") + ylab("Crashes") + 
  theme(title = element_text(size = 9), legend.position = "bottom")
```

#### Multi-Step TBATS (fix)

```{r tbats-ms-stationary-noncovid, echo=FALSE, message=FALSE, include=FALSE}
#Stationary
dif_y2.noncovid <- msts(dif_crashests.noncovid, seasonal.periods=c(7,365.25))

dif_training_TBATS2.noncovid <- subset(dif_y2.noncovid, end=length(dif_y2.noncovid)-426)

dif_test_TBATS2.noncovid <- subset(dif_y2.noncovid, start=length(dif_y2.noncovid)-425)

dif_crashes.train_TBATS2.noncovid <- tbats(dif_training_TBATS2.noncovid)

dif_fc_train_TBATS2.noncovid <- forecast::forecast(dif_crashes.train_TBATS2.noncovid, h=426)

dYhat_ms_TBATS_noncovid <- dif_fc_train_TBATS2.noncovid$mean

y2.noncovid <- msts(crashests.noncovid, seasonal.periods=c(7,365.25))

training_TBATS2.noncovid <- subset(y2.noncovid, end=length(y2.noncovid)-426)

test_TBATS2.noncovid <- subset(y2.noncovid, start=length(y2.noncovid)-425)

Yhat_ms_TBATS_noncovid <- cumsum(c(crashests.noncovid[length(crashests.noncovid)], dYhat_ms_TBATS_noncovid))

Yhat_ms_TBATS_noncovid <- cumsum(c(training_TBATS2.noncovid[length(training_TBATS2.noncovid)], dYhat_ms_TBATS_noncovid)) 

Yhat_ms_TBATS_noncovid <- ts(Yhat_ms_TBATS_noncovid, start = c(2019,1), frequency=365)
```

```{r tbats-ms-stationary-noncovid-plots, echo=FALSE, message=FALSE, warning=FALSE}
#Stationary
autoplot(training_TBATS2.noncovid) + 
  autolayer(Yhat_ms_TBATS_noncovid) + 
  autolayer(test_TBATS2.noncovid) + 
  ggtitle("Multi-Step TBATS Daily Forecasts of Crashes (Stationary Without Pandemic Data)") + 
  xlab("Date") + ylab("Crashes") + 
  theme(legend.position = "bottom")

dif_crashes.test_TBATS2_noncovid <- tbats(dif_test_TBATS2.noncovid)

accuracy(dif_crashes.test_TBATS2_noncovid)

```

#### ARIMA Model 

#### Multi-Step ARIMA 

#### Combination 

### Monthly with Pandemic Data 

#### Stationarity Check 

```{r stationarity-trans-monthly-covid}
##Looking at stationarity of crashes_mts4 monthly time series 
library(urca)
acf(crashes_mts4)
summary(ur.kpss(crashes_mts4))

ndiffs(crashes_mts4) #1
nsdiffs(crashes_mts4) #0

#Attempt at differencing 
dif_crashes_mts4 <- diff(crashes_mts4)

#Looking at stationarity of first difference 
acf(dif_crashes_mts4)
summary(ur.kpss(dif_crashes_mts4))

ndiffs(dif_crashes_mts4) #0
nsdiffs(dif_crashes_mts4) #0
```

```{r stat-trans-monthly-covid-plot}
cbind("Crashes" = crashes_mts4,
      "First differenced" = diff(crashes_mts4)) %>%
  autoplot(facets=TRUE) +
  xlab("Date") + ylab("") +
  ggtitle("Stationarity Transformation of Monthly Pandemic Data")
```

#### TBATS Model 

```{r tbats-stationary-monthly, echo=FALSE, message=FALSE, include=FALSE}
#Stationary

#reload libraries 
crashes_mts = crashes %>%
  filter(as.Date(crash_date) >= "2015/01/01" & as.Date(crash_date) <= "2021/05/31") %>%
  separate(crash_date, 
           into = c("Date", "Hour"), sep = 11) %>%
  group_by(Date) %>%
  summarize(count = length(unique(key_crash))) %>%
  separate(Date, into = c("Year", "Month", "Day"), sep = "/") %>%
  group_by(Year, Month) %>%
  summarise(mcount = sum(count)) %>%
  tidyr::spread(key=Month, value=mcount)

crashes_mts = as.data.frame(crashes_mts)

rownames(crashes_mts) = seq(2015, 2021)

crashes_mts3 = crashes_mts %>%
  select(02:13)

crashes_mts2 <- ts(c(t(crashes_mts3)), frequency=12)

crashes_mts4 <- window(crashes_mts2, start=c(1,01), end=c(7,05), frequency=12)

dif_crashes_mts4 <- diff(crashes_mts4)

dif_y_monthly <- msts(dif_crashes_mts4, seasonal.periods=12)

dif_fit_tbats_covid2_monthly <- tbats(dif_y_monthly)

dif_fc_tbats_covid2_monthly <- forecast::forecast(dif_fit_tbats_covid2, h=8)

dYhat_monthly <- dif_fc_tbats_covid2_monthly$mean #point forecast values 

#crashests2
#dif_crashests2<-diff(crashests2)
#z0<-cumsum(c(crashests2[1], dif_crashests2))
#all(crashests2==z0) #TRUE

#dif_crashests2 <- diff(crashests2) #already ran earlier
#dYhat <- dif_fc_tbats_covid2$mean #already ran earlier 
Yhat_monthly <- cumsum(c(crashes_mts4[length(crashes_mts4)],dYhat_monthly))
Yhat_monthly <- ts(Yhat_monthly, start = c(7,6), frequency=12)
```

```{r tbats-stationary-plot-monthly, echo=FALSE, message=FALSE}
#Stationary
autoplot(crashes_mts4) + 
  autolayer(Yhat_monthly, series="Point Forecasts") + 
  ggtitle("TBATS Forecasting Model for Monthly Car Crashes (With Stationary Pandemic Data)") +
  xlab("Date") + ylab("Crashes") + 
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7, 8), 
                     labels = c("Jan 2015", "Jan 2016", "Jan 2017", "Jan 2018", 
                                "Jan 2019", "Jan 2020", "Jan 2021", "Jan 2022")) + 
  theme(axis.text.x = element_text(angle = 90), title = element_text(size=9), 
        legend.position = "bottom") 
```

#### Multi-Step TBATS

```{r tbats-ms-stationary-monthly, echo=FALSE, message=FALSE, include=FALSE}
#Stationary
dif_y2_monthly <- msts(dif_crashes_mts4, seasonal.periods=12)

dif_training_TBATS2_monthly <- subset(dif_y2_monthly, end=length(dif_y2_monthly)-5)

dif_test_TBATS2_monthly <- subset(dif_y2_monthly, start=length(dif_y2_monthly)-4)

dif_crashes.train_TBATS2_monthly <- tbats(dif_training_TBATS2_monthly)

dif_fc_train_TBATS2_monthly <- forecast(dif_crashes.train_TBATS2_monthly, h=5)

dYhat_ms_TBATS_monthly <- dif_fc_train_TBATS2_monthly$mean

y2 <- msts(crashes_mts4, seasonal.periods=12)

training_TBATS2 <- subset(y2, end=length(y2)-5)

test_TBATS2 <- subset(y2, start=length(y2)-4)
  
#Yhat_ms_TBATS <- cumsum(c(crashests2[length(crashests2)], dYhat_ms_TBATS))
Yhat_ms_TBATS_monthly <- cumsum(c(training_TBATS2[length(training_TBATS2)], dYhat_ms_TBATS_monthly))

Yhat_ms_TBATS_monthly <- ts(Yhat_ms_TBATS_monthly, start = c(6,12), frequency=12)
```

```{r tbats-ms-stationary-plot, echo=FALSE}
#Stationary 
autoplot(training_TBATS2) + 
  autolayer(Yhat_ms_TBATS_monthly, series="Point Forecasts") + 
  autolayer(test_TBATS2, series="Test Set", alpha = 0.7) + 
  ggtitle("Multi-Step TBATS Monthly Forecasts of Crashes (With Stationary Pandemic Data)") + 
  xlab("Date") + ylab("Crashes") + 
  theme(legend.position = "bottom", title=element_text(size=10))

dif_crashes.test_TBATS2_monthly <- tbats(dif_test_TBATS2_monthly)

accuracy(dif_crashes.test_TBATS2_monthly)
```

#### ARIMA Model 

#### Multi-Step ARIMA 

#### HoltWinters Model 

```{r monthly-HW-fcast-decomp-stationary, echo=FALSE}
#Stationary
dif.covid.monthly <- HoltWinters(dif_crashes_mts4)

#summary(dif.covid.monthly)

#plot(fitted(dif.covid.monthly), main = "Box Jenkins Decomposition of Monthly Crashes (With Stationary Pandemic Data)", cex.main=1)

dif.fcast.covid.monthly <- forecast::forecast(dif.covid.monthly, h=8, level=c(80,95))

dYhat.HW <- dif.fcast.covid.monthly$mean

#crashes_mts4
#dif_crashes_mts4<-diff(crashes_mts4)
#z0<-cumsum(c(crashes_mts4[1], dif_crashes_mts4))
#all(crashes_mts4==z0) #TRUE

Yhat.HW <- cumsum(c(crashes_mts4[length(crashes_mts4)],dYhat.HW))
Yhat.HW <- ts(Yhat.HW, start = c(7, 5), frequency=12)
```

```{r monthly-HW-fcast-filter-forecast-ts-stationary, echo=FALSE, message=FALSE}
#Stationary 

autoplot(crashes_mts4) +
  autolayer(Yhat.HW) +
  ggtitle("Forecasts of Monthly Car Crashes Using HoltWinters (With Stationary Pandemic Data)") +
  xlab("Date") + ylab("Crashes") +
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7, 8), 
                     labels = c("Jan 2015", "Jan 2016", "Jan 2017", "Jan 2018", 
                                "Jan 2019", "Jan 2020", "Jan 2021", "Jan 2022")) + 
  theme(axis.text.x = element_text(angle = 90), title = element_text(size=9), 
        legend.position = "bottom")
```

#### Multi-Step HoltWinters 

```{r ms-HW-monthly-forecast-ts-stationary, echo=FALSE, message=FALSE}
#Stationary  
dif_training_HW_monthly <- subset(dif_crashes_mts4, end=length(dif_crashes_mts4)-5)

dif_test_HW_monthly <- subset(dif_crashes_mts4, start=length(dif_crashes_mts4)-4)

dif_crashes.train_HW_monthly <- HoltWinters(dif_training_HW_monthly)

dif_fc_HW <- forecast::forecast(dif_crashes.train_HW_monthly, h=5)

dYhat_ms_HW <- dif_fc_HW$mean

Yhat_ms_HW <- cumsum(c(crashes_mts4[length(crashes_mts4)],dYhat_ms_HW))

Yhat_ms_HW <- ts(Yhat_ms_HW, start = c(7,1), frequency=12)

training_HW_monthly <- subset(crashes_mts4, end=length(crashes_mts4)-5)

test_HW_monthly <- subset(crashes_mts4, start=length(crashes_mts4)-4)

autoplot(training_HW_monthly) + 
  autolayer(Yhat_ms_HW) + 
  autolayer(test_HW_monthly) + 
  ggtitle("Multi-Step HoltWinters Monthly Forecasts of Crashes (With Stationary Pandemic Data)") + 
  xlab("Date") + ylab("Crashes") + 
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7), 
                     labels = c("Jan 2015", "Jan 2016", "Jan 2017", "Jan 2018", 
                                "Jan 2019", "Jan 2020", "Jan 2021")) + 
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom", 
        title = element_text(size = 9))
``` 

#### Combination 

### Monthly without Pandemic Data 

#### Stationarity Check 

```{r stationarity-trans-monthly-noncovid, echo=FALSE, include=FALSE}
##Looking at stationarity of crashes_mts4.noncovid monthly time series 
library(urca)
acf(crashes_mts4.noncovid)
summary(ur.kpss(crashes_mts4.noncovid))

ndiffs(crashes_mts4.noncovid) #1
nsdiffs(crashes_mts4.noncovid) #1

#Attempt at seasonal differencing (noncovid)
season_dif_crashes_mts4.noncovid <- diff(crashes_mts4.noncovid,12)

#Looking at stationarity of seasonal difference (noncovid)
acf(season_dif_crashes_mts4.noncovid)
summary(ur.kpss(season_dif_crashes_mts4.noncovid))

ndiffs(season_dif_crashes_mts4.noncovid) #0
nsdiffs(season_dif_crashes_mts4.noncovid) #0
```

```{r stat-trans-monthly-noncovid-plot, echo=FALSE}
cbind("Crashes" = crashes_mts4.noncovid,
      "Seasonally\n differenced" = diff(crashes_mts4.noncovid, 12)) %>%
  autoplot(facets=TRUE) +
  xlab("Date") + ylab("") +
  ggtitle("Stationarity Transformation of Monthly NonPandemic Data")
```

#### SARIMA Model 

#### Multi-Step SARIMA 

#### STLF Model 

```{r stlf, echo=FALSE, message=FALSE}
q <- crashes_mts4.noncovid %>%
  stlf(lambda = 0, h = 23, level=c(80,95)) 

q %>%
  autoplot() + 
  ggtitle("STLF Model for Monthly Car Crashes (Without Pandemic Data)") +
  xlab("Date") + ylab("Crashes") + 
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8), 
                     labels = c("Jan 2015", "Jan 2016", "Jan 2017", "Jan 2018", 
                                "Jan 2019", "Jan 2020", "Jan 2021", "Jan 2022")) + 
  theme(axis.text.x = element_text(angle = 90))
```

#### Multi-Step STLF 

```{r noncovid-ms-STLF-forecast-ts, echo=FALSE, message=FALSE}
training_STLF_noncovid_monthly <- subset(crashes_mts4.noncovid, end=length(crashes_mts4.noncovid)-14)

test_STLF_noncovid_monthly <- subset(crashes_mts4.noncovid, start=length(crashes_mts4.noncovid)-13)

crashes.train_STLF_noncovid_monthly <- stlf(training_STLF_noncovid_monthly, lambda = 0, h = 14, level=c(80,95))

crashes.train_STLF_noncovid_monthly %>%
  forecast::forecast(h=14) %>%
  autoplot() + autolayer(test_STLF_noncovid_monthly) + 
  ggtitle("Multi-Step STLF Monthly Forecasts of Crashes (Without Pandemic Data)") + 
  xlab("Date") + ylab("Crashes") + 
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7), 
                     labels = c("Jan 2015", "Jan 2016", "Jan 2017", "Jan 2018", 
                                "Jan 2019", "Jan 2020", "Jan 2021")) + 
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom")
```

#### Combination 
